---
title: "Modify this script to test with GitHub"
output: html_document
---

This is an Rmarkdown script we will modify to practice commiting with GitHub. We will reformat some data for the US West Coast Ocean Health Index assessment. Along the way, you can get a little more exposure to:

- Rmarkdown
- reading versioned data directly from github.com
- data wrangling with dplyr, tidyr

## Setup

```{r setup, message=F}

# load packages
library(dplyr)   # install.packages('dplyr')
library(tidyr)   # install.packages('tidyr')
library(knitr)   # install.packages('knitr')

# directories
dir_sandbox = '~/github/github-intro/sandbox'

```


## Explore data read in directly from a different github repository

- navigate to [github.com/OHI-Science/ohi-uswest/scores/scores_index.csv](https://github.com/OHI-Science/ohi-uswest/scores/scores_index.csv)
- click the 'history' button
- click the `<>` button
- click the 'raw' button and copy the url to read directly into R

```{r explore data, message=F}

us_data = read.csv('https://raw.githubusercontent.com/OHI-Science/ohi-uswest/bb2ca4ddc0746edc2f94b5ce07de54e4781f85ce/scores/scores_index.csv')
# us_data

```

## Clean up labeling, remove unneeded columns

```{r clean data}

# relabel
us_data$region_id[us_data$region_id == 8] = 0                   # by convention, study area is region_id 0
us_data$region_label = as.character(us_data$region_label)       # wouldn't rename as factor
us_data$region_label[us_data$region_id == 0] = 'US West Coast'  # simplify name

# remove unneeded columns
us = us_data %>%
  dplyr::arrange(region_id) %>%
  dplyr::select(-region_code, -km2, -preservationist, -extractive, -nonextractive, -extreme, -cc_workshop) %>%
  dplyr::filter(region_label != 'California Current Ocean') %>%
  dplyr::rename(Index = unweighted)

```

## Wrangle

```{r wrangle, eval=F}

# format long and round scores
us = us %>%
  tidyr::gather(goal, score, -region_id, -region_label) %>%
  dplyr::mutate(score = round(score, 2))

```

## Save

```{r save}

write.csv(us, file.path(dir_sandbox, 'scores-ohi-uswest.csv'), row.names =F)

```

